Model for DES Type Ia Supernova Cosmology Analysis

### 2016-05-23 tex/toy.tex
add discussion on evaluating the integral for nice cases of sample selection.

The likelihood is then  re-expressed as
\begin{align}
\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO},
\label{likelihood:eqn}
\end{align}
where  the denominator $p(\text{in analysis}|M, M_I)$ is set
by the condition that the integral over the likelihood over all possible
observations must equal to one.
A frequentist analysis can proceed with this likelihood.
Alternatively, the Bayesian posterior is specified by parameter priors as 
\begin{align}
p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M) \nonumber \\
 & \propto \mathcal{L}_S(O_o| M, L, \tau) p(L| \tau, M) p(\tau|M) p(M).
\label{posterior:eqn}
\end{align}
Once all the $p$'s and $P$'s introduced in
Eqns.~\ref{p_S:eqn}, \ref{likelihood:eqn}, and \ref{posterior:eqn}
are specified we are ready to analyze!

@@

\item $P(anal=True|T=SNIa,c, M, M_I)$:
As an illustrative example we write out the probability of a analysis-sample selection requiring two photometric points above
a signal-to-noise threshold $STON_T$: other sample selection cases are similar but with different combinatorics
\begin{align}
P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)) \nonumber \\
&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
 \end{align}
 where $\mathcal{H}$ is the Heaviside step function.

\item $P(sobs=True|T=SNIa,c, M, M_I)$:
The probability a transient has a spectroscopic observation.
Mat Smith and Chris D'Andrea should be able to provide this term.
Until we get this, lets make the simplifying ansatz that
\begin{align}
P(sobs=True | T=SNIa, c, M, M_I)= & const.
 \end{align}

@@

\subsection{$p(\text{in analysis}|M, M_I)$: The Normalization}

\section{Implementation Notes}
Anticipating the use of available statistical packages, the likelihoods have been constructed
from standard distribution functions.  The truncated distributions from sample selection
are accounted by the normalization 
terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
in the PGM, they are treated as nodes who contribute 
$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
factor.

The following integral appears in the normalization
\begin{equation}
\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
\end{equation}
We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
\begin{equation}
n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
\end{equation}
the integral becomes
 \begin{equation}
\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
\end{equation}
The two probability distributions are normal so that this reduces to
 \begin{equation}
\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
\end{equation}
where
\begin{equation}
c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
\end{equation}

\end{document}

>>>>

The likelihood is then  re-expressed as
\begin{align}
\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\prod_{\alpha \in transients} \int  p_S(O^\alpha |M, M_I)dO^\alpha},
\label{likelihood:eqn}
\end{align}
where  the denominator $p(\text{in analysis}|M, M_I)$ is set
by the condition that the integral over the likelihood over all possible
observations must equal to one.
A frequentist analysis can proceed with this likelihood.
Alternatively, the Bayesian posterior is specified by parameter priors as 
\begin{align}
p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M) \nonumber \\
 & \propto \mathcal{L}_S(O_o| M, L, \tau) p(L| \tau, M) p(\tau|M) p(M).
\label{posterior:eqn}
\end{align}
Once all the $p$'s and $P$'s introduced in
Eqns.~\ref{p_S:eqn}, \ref{likelihood:eqn}, and \ref{posterior:eqn}
are specified we are ready to analyze!

@@

\item $P(anal=True|T=SNIa,c, M, M_I)$:
As an illustrative example we write out the probability of a analysis-sample selection requiring two photometric points above
a signal-to-noise threshold $STON_T$: other sample selection cases are similar but with different combinatorics
\begin{align}
P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)) \nonumber \\
&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
\label{anal:eqn}
 \end{align}
 where $\mathcal{H}$ is the Heaviside step function.

\item $P(sobs=True|T=SNIa,c, M, M_I)$:
The probability a transient has a spectroscopic observation.
Mat Smith and Chris D'Andrea should be able to provide this term.
Until we get this, lets make the simplifying ansatz that
\begin{align}
P(sobs=True | T=SNIa, c, M, M_I)= & 1.
\label{sobs:eqn}
 \end{align}

@@

\subsection{$p(\text{in analysis}|M, M_I)$: The Normalization}
The contribution to the normalization from transient $\alpha$ can be rewritten as
\begin{align}
\int p_S(O^\alpha|M, M_I)dO^\alpha &= I_c I_s P(T^\alpha=SNIa| M, M_I),
\end{align}
where the independence of the photometric and spectroscopic observations allow us to separate the contributions of each
\begin{multline}
I_c = \int  P(disc=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) \\ P(sobs=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha)
p(c^\alpha|M, M_I,z_{host}=z^\alpha) dc\\
\end{multline}
and
\begin{equation}
I_s = \int P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I,z_{host}=z^\alpha)
 p(s^\alpha|M, M_I,z_{host}=z^\alpha)ds.
\end{equation}

Note that in the pure SN~Ia sample analysis, $P(T^\alpha=SNIa| M, M_I)$ divides out in the calculation of the likelihood,  so that the cosmological results
are independent of rates.  This would not be the case for the photometric-sample analysis.

The integrals $I_c$ and $I_s$ can be simple to evaluate in certain circumstances.  Suppose the photometric selection requirements
are set so that $P(disc=True | T=SNIa,c,  M, M_I) \sim 1$ and the photometric uncertainties are independent (approximatively true when
each point's noise is dominated by its image noise).  Then for the case of the sample selection (spectroscopic observation)
criteria in Eqn.~\ref{anal:eqn} (\ref{sobs:eqn}), 
\begin{align}
I_c & = \int  P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) p(c^\alpha|M, M_I,z_{host}=z^\alpha) dc\\
& = 1 - \prod_{\beta \in images} \text{CDF}\left(N_\beta STON_T; \mathcal{N}(c_\star,\sigma_\beta)\right)   \nonumber \\
& - \sum_{\alpha \in images}  \text{CCDF}\left(N_\alpha STON_T;\mathcal{N}(c_\star,\sigma_\alpha)\right)   \prod_{\beta \in \{images-\alpha\}} \text{CDF}\left(N_\beta STON_T;\mathcal{N}(c_\star,\sigma_\beta)\right),
\end{align}
where CDF and CCDF are the cumulative and the complimentary cumulative distribution functions respectively of the normal distribution
that are readily accessible in programming language libraries either on their own or as error fuctions.
%
%\section{Implementation Notes}
%Anticipating the use of available statistical packages, the likelihoods have been constructed
%from standard distribution functions.  The truncated distributions from sample selection
%are accounted by the normalization 
%terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
%in the PGM, they are treated as nodes who contribute 
%$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
%factor.
%
%
%The following integral appears in the normalization
%\begin{equation}
%\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
%\end{equation}
%We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
%\begin{equation}
%n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
%\end{equation}
%the integral becomes
% \begin{equation}
%\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
%\end{equation}
%The two probability distributions are normal so that this reduces to
% \begin{equation}
%\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
%\end{equation}
%where
%\begin{equation}
%c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
%\end{equation}
%
\end{document}
